State of the Practice for Lattice Boltzmann Method Software

### 2022-04-25 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Revisions to Section 8

	\item[P\refstepcounter{pnum}\thepnum \label{P_LackDevTime}:] \textbf{Lack of
	Development Time} A developer of pyLBM noted that their small development
	team has a lack of time to implement new features. Small development teams
	are common for LBM software packages (as shown in the measurement table
	excerpt in Figure~\ref{measurement_template_image}). Other domains of
	research software also experience the lack of time pain point
	\citep{PintoEtAl2018, PintoEtAl2016, WieseEtAl2019}.

	\item[P\refstepcounter{pnum}\thepnum \label{P_LackSoftDevExp}:] \textbf{Lack
	of Software Development Experience} The developer of TCLB noted a lack of
	software development experience, and others noted a need for improving
	software engineering education. Many of the team members are domain experts,
	not computer scientists or software engineers. \citet{Nguyen-HoanEtAl2010}
	noted this same trend, with only 23\% of research software survey
	respondents having a computing-related background. Similarly,
	\citet{UditAndKatz2017} show that the majority (54\%) of postdocs have not
	received training in software development.  The LBM developer suggesting an
	increasing role for formal software education matches the trend observed by
	\citet{PintoEtAl2018}, where their replication of a previous study
	\citep{HannayEtAl2009}, shows a growing interest in formal training (From
	13\% of respondents in 2009 to 22\% in 2018). \citet{PintoEtAl2018} found
	that some developers feel there is a mismatch between coding skills and
	subject-matter skills.

	\item[P\refstepcounter{pnum}\thepnum \label{P_LackFunding}:] \textbf{Lack of
	Incentive and Funding} The TCLB developer noted a lack of incentives and
	funding in academia for developing widely used research software. This
	problem has also been noted by others \citep{gewaltig2012quality, Goble2014,
	KaterbowAndFeulner2018}.  \citet{WieseEtAl2019} reported developer pains
	related to publicity, since historically publishing norms make it difficult
	to get credit for creating software.  As studied by
	\citet{HowisonAndBullard2016}, research software (specifically biology
	software, but the trend likely applies to other research software domains)
	is infrequently cited.  \citet{PintoEtAl2018} also mentions the lack of
	formal reward system for research software.

	\item[P\refstepcounter{pnum}\thepnum \label{P_LackExtSupport}:] \textbf{Lack
	of External Support} Interviewees raised a concern that there are no
	organizations helping with the development of good quality software.  The
	literature on research software does not tend to echo this concern because
	there are such organizations, including \href{https://bssw.io/} {Better
	Scientific Software (BSSw)},
	\href{https://www.software.ac.uk/} {Software Sustainability Institute}
	\citep{CrouchEtAl2013}, and \href{https://software-carpentry.org/}{Software
	Carpentry} \citep{WilsonAndLumsdaine2006, Wilson2016}. Over time awareness
	of these groups will grow, so this pain point is likely to disappear in the
	future for LBM and other research software developers.

	\item[P\refstepcounter{pnum}\thepnum \label{P_TechnologyHurdles}:]
	\textbf{Technology Hurdles} Technology pain points for LBM developers
	include setting up parallelization and CI. The pain
	point survey of \citet{WieseEtAl2019} also highlighted technical-related
	problems like dependency management, cross-platform compatibility,
	CI, hardware issues and operating system issues.

	\item[P\refstepcounter{pnum}\thepnum \label{P_Correctness}:]
	\textbf{Ensuring Correctness} Several developers noted difficulties with
	ensuring correctness. They alluded to difficulty with testing the
	correctness of large numbers of features, and challenges with automated
	testing. The TCLB developer commented that the amount of testing data needed
	was sometimes problematic, since free testing services do not offer adequate
	facilities for large amounts of data, requiring in-house testing solutions.
	Other research software domains point to the following problems with
	testing: i) \citep{PintoEtAl2018} mention the problem of insufficient
	testing; ii) the survey of \citep{HannayEtAl2009} shows that more developers
	think testing is important than the number that believe they have a
	sufficient understanding of testing concepts; and, iii) previous studies
	\citep{HannayEtAl2009, KanewalaAndBieman2013, KellyEtAl2011, WieseEtAl2019}
	point to the oracle problem, which occurs in research software when we do
	not have a means to judge the correctness of the calculated solutions.  The
	LBM experience seems to overlap with i and ii, but the LBM developers did
	not allude to problem iii (the oracle problem) in our conversations.  As
	discussed in Section~\ref{SecReproducibility}, the oracle problem likely did
	not come up because the LBM developers we interviewed have strategies for
	verifying their work.

	\item[P\refstepcounter{pnum}\thepnum \label{P_Usability}:]
	\textbf{Usability} Several developers noted that users sometimes try to use
	incorrect LBM method combinations to solve their problems. Furthermore, some
	users think that the packages will work out of the box to solve their cases,
	while in reality CFD knowledge needs to be applied to correctly modify the
	packages for the new endeavour.  Some respondents in the survey of
	\citet{WieseEtAl2019} also mentioned that users do not always have the
	expertise required to install or use the software.

	\item[P\refstepcounter{pnum}\thepnum \label{P_TechDebt}:] \textbf{Technical
	Debt} The developer of ESPResSo said that they originally wrote their source
	code with a specific application in mind, which later caused too much
	coupling between components in the source code. This results in technical
	debt \citep{KruchtenEtAl2012}, which has an impact on future maintainability
	and reusability. Concern with technical debt is likely why researchers in
	the survey of \citet{Nguyen-HoanEtAl2010} rated maintainability as the third
	most important software quality. More recently the push for sustainable
	software \citep{deSouzaEtAl2019} is motivated by the pain that past
	developers have had with accumulating too much technical debt.

	\item[P\refstepcounter{pnum}\thepnum \label{P_Documentation}:]
	\textbf{Quality of Documentation}  Interviewees stressed the importance of
	documentation for both users and developers throughout the interviews.
	However, it was noted several times that a lack of time and funding
	(\ppref{P_LackDevTime}) has a negative effect on the documentation. Most of
	the developers are scientific researchers evaluated on the scientific papers
	that they produce. Writing and updating documentation is something that is
	done in their free time, if that time arises. Others also mention inadequate
	research software documentation \citep{PintoEtAl2018, WieseEtAl2019}.  The
	problem also arises with non-research software \citep{LethbridgeEtAl2003}.
	Recommendations on the development of research software state that
	developers should critically evaluate their own development processes in
	terms of quality assurance and comply with international standards for
	software documentation \citep{KaterbowAndFeulner2018}.

>>>>

	\item[P\refstepcounter{pnum}\thepnum \label{P_LackDevTime}:] \textbf{Lack of
	Development Time:} A developer of pyLBM noted that their small development
	team has a lack of time to implement new features. Small development teams
	are common for LBM software packages (as shown in the measurement table
	excerpt in Figure~\ref{measurement_template_image}). Other domains of
	research software also experience the lack of time pain point
	\citep{PintoEtAl2018, PintoEtAl2016, WieseEtAl2019}.

	\item[P\refstepcounter{pnum}\thepnum \label{P_LackSoftDevExp}:] \textbf{Lack
	of Software Development Experience:} The developer of TCLB noted a lack of
	software development experience, and others noted a need for improving
	software engineering education. Many of the team members are domain experts,
	not computer scientists or software engineers. \citet{Nguyen-HoanEtAl2010}
	noted this same trend with research software developers, with only 23\% of
	survey respondents having a computing-related background. Similarly,
	\citet{UditAndKatz2017} show that the majority (54\%) of postdocs have not
	received training in software development.  The LBM developer suggesting an
	increasing role for formal software education matches the trend observed by
	\citet{PintoEtAl2018}, where their replication of a previous study
	\citep{HannayEtAl2009}, shows a growing interest in formal training (From
	13\% of respondents in 2009 to 22\% in 2018). \citet{PintoEtAl2018} found
	that some developers feel there is a mismatch between coding skills and
	subject-matter skills.

	\item[P\refstepcounter{pnum}\thepnum \label{P_LackFunding}:] \textbf{Lack of
	Incentive and Funding:} The TCLB developer noted a lack of incentives and
	funding in academia for developing widely used research software. This
	problem has also been noted by others \citep{gewaltig2012quality, Goble2014,
	KaterbowAndFeulner2018}.  \citet{WieseEtAl2019} reported developer pains
	related to publicity, since publishing norms have historically made it
	difficult to get credit for creating software.  As studied by
	\citet{HowisonAndBullard2016}, research software (specifically biology
	software, but the trend likely applies to other research software domains)
	is infrequently cited.  \citet{PintoEtAl2018} also mentions the lack of
	formal reward system for research software.

	\item[P\refstepcounter{pnum}\thepnum \label{P_LackExtSupport}:] \textbf{Lack
	of External Support:} Interviewees raised a concern that there are no
	organizations helping with the development of good quality software.  The
	literature on research software does not echo this concern because
	there are such organizations, including \href{https://bssw.io/} {Better
	Scientific Software (BSSw)},
	\href{https://www.software.ac.uk/} {Software Sustainability Institute}
	\citep{CrouchEtAl2013}, and \href{https://software-carpentry.org/}{Software
	Carpentry} \citep{WilsonAndLumsdaine2006, Wilson2016}. Over time awareness
	of these groups will grow, so this pain point is likely to disappear in the
	future for LBM and other research software developers.

	\item[P\refstepcounter{pnum}\thepnum \label{P_TechnologyHurdles}:]
	\textbf{Technology Hurdles:} Technology pain points for LBM developers
	include setting up parallelization and CI. The pain
	point survey of \citet{WieseEtAl2019} also highlighted technical-related
	problems like dependency management, cross-platform compatibility,
	CI, hardware issues and operating system issues.

	\item[P\refstepcounter{pnum}\thepnum \label{P_Correctness}:]
	\textbf{Ensuring Correctness:} Several developers noted difficulties with
	ensuring correctness. They alluded to difficulty with testing the
	correctness of large numbers of features, and challenges with automated
	testing. The TCLB developer commented that the amount of testing data needed
	was sometimes problematic, since free testing services do not offer adequate
	facilities for large amounts of data, requiring in-house testing solutions.
	Other research software domains point to the following problems with
	testing: i) \citet{PintoEtAl2018} mention the problem of insufficient
	testing; ii) \citet{HannayEtAl2009} show that more developers
	think testing is important than the number that believe they have a
	sufficient understanding of testing concepts; and, iii) 
	\citet{HannayEtAl2009, KanewalaAndBieman2013, KellyEtAl2011, WieseEtAl2019}
	point to the oracle problem, which occurs in research software when we do
	not have a means to judge the correctness of the calculated solutions. Based
	on the interviews, the LBM experience seems to overlap with i and ii; the
	developers did not allude to problem iii (the oracle problem).  As discussed
	in Section~\ref{SecReproducibility}, the oracle problem likely did not come
	up because the LBM developers have strategies for verifying their work.

	\item[P\refstepcounter{pnum}\thepnum \label{P_Usability}:]
	\textbf{Usability:} Several developers noted that users sometimes try to use
	incorrect LBM method combinations to solve their problems. Furthermore, some
	users think that the packages will work out of the box to solve their cases,
	while in reality CFD knowledge needs to be applied to correctly modify the
	packages for the new endeavour.  Some respondents in the survey of
	\citet{WieseEtAl2019} also mentioned that users do not always have the
	expertise required to install or use the software.

	\item[P\refstepcounter{pnum}\thepnum \label{P_TechDebt}:] \textbf{Technical
	Debt:} The developer of ESPResSo said that they originally wrote their source
	code with a specific application in mind, which later caused too much
	coupling between components in the source code. This resulted in technical
	debt \citep{KruchtenEtAl2012}, which has an impact on future maintainability
	and reusability. Concern with technical debt is likely why researchers in
	the survey of \citet{Nguyen-HoanEtAl2010} rate maintainability as the third
	most important software quality. More recently the push for sustainable
	software \citep{deSouzaEtAl2019} is motivated by the pain that past
	developers have had with accumulating too much technical debt.

	\item[P\refstepcounter{pnum}\thepnum \label{P_Documentation}:]
	\textbf{Quality of Documentation:}  Interviewees stressed the importance of
	documentation for both users and developers throughout the interviews. They
	emphasize that a lack of time and funding (\ppref{P_LackDevTime}) has a
	negative effect on the documentation. Most of the developers are scientific
	researchers evaluated on the scientific papers that they produce. Writing
	and updating documentation is something that is done in their free time, if
	that time arises. Others also mention inadequate research software
	documentation \citep{PintoEtAl2018, WieseEtAl2019}.  The problem also arises
	with non-research software \citep{LethbridgeEtAl2003}. Recommendations on
	the development of research software state that developers should critically
	evaluate their own development processes in terms of quality assurance and
	comply with international standards for software documentation
	\citep{KaterbowAndFeulner2018}.
